Manuscript 1 code

### 2024-09-11 manuscript_1_code.Rmd
Edits to Fig1 code

functional_group_colours <- c("coral" = "coral",
                         "cca" = "pink",
                         "maroalgae" = "forestgreen",
                         "turf" =  "darkolivegreen")

@@

simplified_cover <- data.frame(
  year = cover_df$year,
  habitat = cover_df$habitat,
  site = cover_df$site,
  location = cover_df$location,
  low_lying_algae = cover_df$cca + cover_df$turf + cover_df$cyanophyta,
  coral = cover_df$coral,
  macroalgae = cover_df$macroalgae
)

simplified_cover_means <- simplified_cover %>%
  group_by(year, habitat) %>%
  summarise(macroalgae_mean = mean(macroalgae),
            low_lying_algae_mean = mean(low_lying_algae),
            coral_mean = mean(coral),
            macroalgae_se = sd(macroalgae)/sqrt(n()),
            low_lying_algae_se = sd(low_lying_algae)/sqrt(n()),
            coral_se = sd(coral)/sqrt(n())
            )

simplified_cover_long <- pivot_longer(
  data = simplified_cover_means,
  cols = 3:5,
  names_to = "functional_group",
  values_to = "percent_cover"
)

simplified_cover_long_se <- pivot_longer(
  data = simplified_cover_means,
  cols = 6:8,
  names_to = "functional_group",
  values_to = "se"
)
simplified_cover_long$se <- simplified_cover_long_se$se
```

@@

```{r figure 1, fig.height = 10, fig.width=14}
(figure_1 <- ggplot(data = subset(cover_long, functional_group != "sand_mean" & functional_group != "other_mean")) +
  geom_line( aes(x = year, y = percent_cover, colour = functional_group), size = 1) +
  geom_ribbon( aes(x = year, ymin = percent_cover - se, ymax = percent_cover + se, fill = functional_group), alpha = 0.3) +
  facet_wrap(~ habitat) +
  geom_segment(aes(y = 75, yend = 75, x = 2006, xend = 2010), linewidth = 2, colour = "grey") +
#  scale_colour_manual(values = functional_group_colours) +
#  scale_fill_manual(values = functional_group_colours) +
  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black")) )

>>>>

functional_group_colours <- c("coral_mean" = "#56B4E9",
                         "cca_mean" = "#CC79A7",
                         "macroalgae_mean" = "#009E73",
                         "turf" =  "#D55E00")

@@

cover_long$habitat <- factor(cover_long$habitat, 
                             levels = c("fringing", "backreef", "outer_10","outer_17"),
                             labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))

# simplified_cover <- data.frame(
#   year = cover_df$year,
#   habitat = cover_df$habitat,
#   site = cover_df$site,
#   location = cover_df$location,
#   low_lying_algae = cover_df$cca + cover_df$turf + cover_df$cyanophyta,
#   coral = cover_df$coral,
#   macroalgae = cover_df$macroalgae
# )
# 
# simplified_cover_means <- simplified_cover %>%
#   group_by(year, habitat) %>%
#   summarise(macroalgae_mean = mean(macroalgae),
#             low_lying_algae_mean = mean(low_lying_algae),
#             coral_mean = mean(coral),
#             macroalgae_se = sd(macroalgae)/sqrt(n()),
#             low_lying_algae_se = sd(low_lying_algae)/sqrt(n()),
#             coral_se = sd(coral)/sqrt(n())
#             )
# 
# simplified_cover_long <- pivot_longer(
#   data = simplified_cover_means,
#   cols = 3:5,
#   names_to = "functional_group",
#   values_to = "percent_cover"
# )
# 
# simplified_cover_long_se <- pivot_longer(
#   data = simplified_cover_means,
#   cols = 6:8,
#   names_to = "functional_group",
#   values_to = "se"
# )
# simplified_cover_long$se <- simplified_cover_long_se$se

@@

```{r figure 1, fig.height = 10, fig.width=14}
(figure_1 <- ggplot(data = subset(cover_long, functional_group != "sand_mean" & functional_group != "other_mean")) +
  geom_line(aes(x = year, y = percent_cover, colour = functional_group), linewidth = 1) +
  geom_ribbon( aes(x = year, ymin = percent_cover - se, ymax = percent_cover + se, fill = functional_group), alpha = 0.3) +
  facet_wrap(~ habitat) +
  geom_segment(aes(y = 75, yend = 75, x = 2006, xend = 2010), linewidth = 2, colour = "grey") +
  scale_colour_manual(values = functional_group_colours) +
  scale_fill_manual(values = functional_group_colours) +
  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
   scale_x_continuous(breaks = unique(cover_long$year)) +
   labs(y = "Percent Cover \n",
        x = "Year") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black")) )

ggsave(filename = "output/figure_1.png", figure_1, height = 10, width = 14)
